---
title: "1-Solutions-GenDistance"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Users/ravengriffin/Downloads/Evolution")

### Load packages (install if you haven't already)
library(msa)
library(phangorn)
library(phytools)
```

# Part I: Read in and Align the Sequences with Clustal-Omega
```{r alignment}
#Align sequences
dnaSeqs = readDNAStringSet("flu_seqs.fasta")
aligns = msa(dnaSeqs, method="ClustalOmega", order="input")


# ── Load & normalize ──────────────────────────────────────────────────────────
```

# Part II: Use the Phangorn package to compare substitution models
```{r alignment}
forPhang = msaConvert(aligns, type="phangorn::phyDat")

mt = modelTest(forPhang, model=c("JC","F81","K80","HKY","TrN"))
write.table(mt, "modelTest_results.txt", quote=FALSE, sep="\t",row.names=FALSE, col.names=TRUE)
```

## Question 1: The Bayesian Information Criterion (BIC) is a method for comparing models that includes penalties for more complex models. With this method, the best one is the model with the LOWEST BIC score. In your test, which model is this? Describe this model in terms of its assumptions. (5 pts)

YOUR ANSWER TO QUESTION ONE HERE.

# Part III: Calculate Genetic Distance Under the Best Model

## Question 2: Show your command for calculating distance, including which model you used in your command, and which option you used for gamma. (1 pt)

```{r gendist}
dna = as.DNAbin(forPhang)
?dist.dna()
D = dist.dna(dna, model="TN93", gamma=FALSE)
```

# Part IV: Create both UPGMA and Neighbor-Joining Trees
```{r trees, fig.dim=c(8,8)}
t.upgma = upgma(D)
t.nj = nj(D)
t.upgma.root = root(phy=t.upgma, outgroup='1993a')
t.nj.root = root(phy=t.nj, outgroup='1993a')
plot(t.upgma.root, main="UPGMA", cex=0.75)
plot(t.nj.root, main="Neighbor-Joining", cex=0.75)
treedist(t.upgma.root, t.nj.root)
```

## Question 3: Provide your plots for the rooted UPGMA and rooted Neighbor-Joining trees (be sure they are labeled so I can tell which is which!). Describe any differences you see in the overall appearance of the two trees. Do the same clades appear in each tree? Do viruses collected in later years branch off from the viruses of previous years (like we would expect)? (5 pts)

YOUR ANSWER TO QUESTION 3 HERE.

# Part V: Calculate Distance Metrics Between the 2 Trees

## Question 4: What is the symmetric difference, branch score difference, and path difference between your two trees?  Which number implies the biggest difference in the trees?  Based on what you know about the definition of these metrics, why would one score be much higher than the others?  Does your highest score make sense given what you see when you plot the two trees? (3 pts)

YOUR ANSWER TO QUESTION 4.

# Part VI: Perform Bootstrapping on the Neighbor-Joining Tree

## Question 5: Provide a plot of your bootstrapped tree. Are there any nodes in your tree that seem to have very weak bootstrap support (i.e. less than 50%)? How many of these weakly supported nodes do you see? Which node has the weakest support, and what clade is this node at the base of? If it is hard to see exactly where the nodes are with the bootstrap labels on them, you might also want to refer back to your original plot without the labels to help you. (3 pts)

```{r bootstrap, fig.dim=c(8,8)}
myBoots = boot.phylo(t.nj.root, dna, function(x) root(nj(dist.dna(x, model="TN93", gamma=TRUE)),1), trees=TRUE)
plot(t.nj.root, cex=0.5, edge.width=2, main="Some Title for Your Tree")
nodelabels(myBoots$BP, cex=0.4)
```
YOUR ANSWER TO QUESTION 5.

## Question 6: Provide a plot of your majority rule consensus tree. (1 pt)
```{r majorityrule, fig.dim=c(8,8)}
mjtree = consensus(myBoots$trees, p=0.5)
mjtree.root = root(mjtree, outgroup='1993a')
plot(mjtree.root, cex=0.5, edge.width=2, main="Majority Rule Consensus Tree")
```

# Part VII: Find the Most Parsimonious Tree

## Question 7: What are your parsimony scores for your UPGMA and Neighbor-Joining trees? Given that a lower score is better, which of your trees is the most parsimonious? Based on what you know about the definition of parsimony, what does it mean when we say that one tree is more parsimonious than another? i.e. How should we interpret this result? (3 pts)
```{r parsimony}
parsimony(t.upgma.root, forPhang)
parsimony(t.nj.root, forPhang)
tre.pars = pratchet(forPhang, start=t.nj)

```
YOUR ANSWER TO QUESTION 7.

## Question 8: Were you able to improve on the parsimony score you got with the neighbor-joining tree? (1 pt)
YOUR ANSWER TO QUESTION 8.

## Question 9: Provide the plot for your final tree. Even though there are still some uncertain relationships, can you now see any kind of trend with respect to the years the viruses were sampled and where they appear on the tree? Describe your new tree and the trends that you see. (3 pts)

```{r final, fig.dim=c(8,8)}
N <- length(t.nj.root$tip.label)
toCollapse <- match(which(myBoots$BP<50)+N, t.nj.root$edge[,2])
t.nj.root$edge.length[toCollapse] = 0
new.tree = di2multi(t.nj.root, tol=0.00001)
plot(new.tree, type="unr", show.tip=FALSE, edge.width=2)
title("Maximum-parsimony tree")
install.packages("basetheme")
install.packages("epicontacts")
library(basetheme)
library(epicontacts)
year = as.numeric(gsub("[a-c]", "", names(forPhang)))
tiplabels(t.nj.root$tip.label, bg=transp(num2col(year),.7), cex=.5, fg="transparent")
```

YOUR ANSWER TO QUESTION 9.